---
title: "SaudiOilStockAnalysis"
author: "Creig"
date: "2024-10-01"
output: html_document
---

Importing the data file 'SaudiArabia.xlsx' that contains data related to Saudi Arabia stock market and oil market.

```{r}
library(readxl)
library(tseries)
library(ggplot2)
library(zoo)


stockdata<- read_excel("D:\\Master's Files\\Time Series\\SaudiArabia.xlsx", sheet = 'Tadawul Share')
stockdata
```

```{r}
summary(stockdata)

```

Getting the dates with the lowest and highest closing prices.
```{r}

paste(stockdata$Date[stockdata$Price==min(stockdata$Price)])

paste(stockdata$Date[stockdata$Price==max(stockdata$Price)])
```


Getting the dates with the lowest and highest 'High' prices.
```{r}

paste(stockdata$Date[stockdata$High==min(stockdata$High)])

paste(stockdata$Date[stockdata$High==max(stockdata$High)])
```


Getting the dates with the lowest and highest closing prices.
```{r}

paste(stockdata$Date[stockdata$`Change %`==min(stockdata$`Change %`)])

paste(stockdata$Date[stockdata$`Change %`==max(stockdata$`Change %`)])
```


Converting the data frame to 'zoo' format.
```{r}

stockzoo<-read.zoo(data.frame(stockdata),FUN=as.Date, format="%d%m%y")

print("Beginning Observations:")
head(stockzoo)
print("End Observations:")
tail(stockzoo)

```

Plotting the closing price for Saudi Arabia's stock market as time series.

```{r}
library(ggplot2)
library(plotly)

attach(stockdata)
ggplot(data=stockdata, aes(x=Date, y=Price))+geom_line()
```

Plotting the lowest price of the day.

```{r}

attach(stockdata)
ggplot(data=stockdata, aes(x=Date,y=Low))+geom_area()

```

Plotting all data points
```{r}
windows.options(width=200, height=600)
#dev.new(width=10, height=5, unit="in")
plot(stockzoo)

```


plotting the open price and the date
```{r}

ggplot(stockdata, aes(Date,Open))+geom_line()


```

Plotting the Highest price and the date
```{r}

ggplot(stockdata, aes(Date,High))+geom_line()


```

Plotting the change as a percentage.

```{r}

attach(stockdata)

stockdata$ChangeNum<-as.numeric(stockdata$`Change %`)

  
windows.options(height=400, width=600)
   
ggplot(stockdata, aes(Date,ChangeNum))+geom_line()


```


ADF Test to check for the stationarity.
```{r}
adf.test(stockdata$Price)

```

ACF plot for the closing price.
```{r}
acf(stockdata$Price)

```


```{r}
cor(stockdata$Price,stockdata$High)

```

Taking first order difference for the close price and high price.
```{r warning=FALSE}
#attach(stockzoo)

retclose<-diff(log(stockdata$Price))
rethigh<-diff(log(stockdata$High))

```


```{r}
acf(rethigh)

```

ACF plot for 'retclose'
```{r}
acf(retclose)
```

partial ACF plot for retclose
```{r}
pacf(retclose)

```


Since there is still autocorrelation, we take second order difference of the values.

```{r}
ret2close<-diff(retclose)

```

```{r}
adf.test(ret2close)

```


```{r}
acf(ret2close)

```

```{r}
pacf(ret2close)

```

```{r}
cor(retclose, rethigh)
```


```{r}
adf.test(retclose)
```



Importing the oil data for OPEC Basket.
```{r}

oildata<- read_excel("D:\\Master's Files\\Time Series\\SaudiArabia.xlsx", sheet = 'OPEC Basket')

#oildata$Value<-as.numeric(oildata$Value)

head(oildata)

```


Getting the dates with the lowest and highest closing prices.
```{r}
summary(oildata)

paste(oildata$Date[oildata$Value==min(oildata$Value)])

paste(oildata$Date[oildata$Value==max(oildata$Value)])

```



Converting the oil data to 'zoo' format.
```{r}
oilzoo<- read.zoo(data.frame(oildata), FUN = as.Date, format="%d%m%y")

tail(oilzoo)

summary(oilzoo)

```


Plotting for the Basket value as a time series plot.
```{r}

ggplot(oildata, aes(Date,Value))+geom_line()

```

ACF Test for the Oil Data value.
```{r}
acf(oildata$Value)

```

Taking the first order difference since the data does not show stationarity.
```{r}
retoil<-diff(log(oildata$Value))

acf(retoil)

```


```{r}
pacf(retoil)
```

Taking second order difference of the data
```{r}
ret2oil<-diff(retoil)

pacf(ret2oil)

```


```{r}
acf(ret2oil)

```


```{r}
adf.test(ret2oil)

```


```{r}

saudirawdata <- read_excel("D:\\Master's Files\\Time Series\\SaudiArabia.xlsx", sheet = 'Overall')

tail(saudirawdata)
```

```{r}

saudidata <- na.omit(saudirawdata)

View(saudidata)

```


```{r}

adf.test(oilzoo)

```



Taking the differences for oil, close and high for the consolidated data.
```{r}

head(oilzoo)

retoil<-diff(log(saudidata$BasketValue))

adf.test(retoil)

```

```{r}
retclose <- diff(log(saudidata$StockPrice))

rethigh <- diff(log(saudidata$High))




adf.test(retclose)

adf.test(rethigh)



```


```{r}

library(lmtest)
attach(oildata)

cor(saudidata$StockPrice,saudidata$BasketValue)


cor(saudidata$StockPrice,saudidata$High)
```

```{r}
ret2oil<-diff(retoil)
ret2close<-diff(retclose)

```


```{r}

options(scipen=999)

vardata <- ts(cbind(retoil, retclose))  


head(vardata)

```


```{r}
library(vars)

lag_selectionX <- VARselect(vardata, lag.max = 10, type = "const")  # Choose maximum lags
print(lag_selectionX)
```




```{r}
var_modelX <- VAR( vardata, p = lag_selection2$selection[1], type = "const")
summary(var_modelX)

```


```{r}
irf(var_modelX, impulse = "retoil", response = "retclose")

```


```{r}
varlatest<- head(vardata, 400)

View(varlatest)

varolder <- tail(vardata, 400)

View(varolder)
```






```{r}
lag_selection1<- VARselect(varlatest, lag.max = 10, type = "const")  # Choose maximum lags
print(lag_selection1)
```



```{r}
var_model1 <- VAR( varlatest, p = lag_selection1$selection[1], type = "const")
var_model1
#summary(var_model1)

```


```{r}
lag_selection2 <- VARselect(varolder, lag.max = 10, type = "const")  # Choose maximum lags
print(lag_selection2)
```




```{r}
var_model2 <- VAR( varolder, p = lag_selection2$selection[1], type = "const")
summary(var_model2)

```



```{r}
irf(var_model1, impulse = "retoil", response = "retclose")

```

```{r}
irf(var_model2, impulse = "retoil", response = "retclose")

```

Transfer Function Model for the whole data.

Here, we are comparing the variable 'close' with the OPEC 'oil' price index.




```{r}
library(forecast)
tf_model <- Arima(ret2close, xreg = ret2oil)
summary(tf_model)

```


```{r}
newret2oil  <-head(ret2oil,400)
newret2close<-head(ret2close,400)

tf_model_new <- Arima(newret2close, xreg = newret2oil)
summary(tf_model_new)

```


```{r}
oldret2oil  <-tail(ret2oil,400)
oldret2close<-tail(ret2close,400)

tf_model_old <- Arima(oldret2close, xreg = oldret2oil)
summary(tf_model_old)

```













